Accelerating Traffic Simulation and Its Real-Time Challenges

### 2012-09-12 draft.tex
added Evaluation

\begin{abstract}
\end{abstract}

@@

\section{Assumption}
\label{sec:assumption}

grid, block, thread.

A unit of functions that is launched on the GPU is called a \textit{kernel}.

@@

@@

Figure~\ref{fig:sample-in-cpu} shows a brief overview of the
Sample-in-CPU scheme.
This scheme also has several alternatives depending on how many kernels
need to be launched on the GPU in each period. 
Suppose that we want to offload $n$ pieces of \textit{for} loops on to
the GPU.
We may return to the CPU $n$ times in totall, that is, return every time
one \textit{for} loop breaks, or otherwise we may just return once when
all the $n$ \textit{for} loops end.
This is a design decision, and is also dependent on the GPU architecture
and the parallelization structure.
In general, parallelized threads need to be synchronized when moving
across basic blocks.
Specifically, when moving to the next \textit{for} loop executed in
parallel, all the threads relevant to this parallel computing procedure
must synchronize with each other.
However, the maximum number of threads that can be synchronized on the
GPU is often limited.
As of 2012, for example, NVIDIA's GPU architectures limit the number of
such threads to 1024 or less.
Our implementation therefore forces the program to return to the CPU
every time one \textit{for} loop breaks.
Note that this is not a conceptual limitation of GPU computing, but is a
current limitation of hardware.
We believe that this limitation would be removed or mitigated in
next-generation GPU architectures.

@@

We evaluate the performance of our GPU-accelerated traffic simulation
programs, using an NVIDIA GeForce GTX 560 Ti graphics card.
This is a middle-end graphics card integrating 394 compute cores on a
chip.

\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{eps/eval_accel.eps}
\caption{Acceleration of traffic simulation.}
\label{fig:eval_accel}
\end{figure}

\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{eps/eval_nosync.eps}
\caption{Impact of synchronization.}
\label{fig:eval_nosync}
\end{figure}

@@

\subsection{Experimental Results}
\label{sec:experimental_results}

@@

Imprecise Computation.
Gdev~\cite{Kato2012}.

>>>>

\begin{abstract}
PLACEHOLDER.
\end{abstract}

@@

The rest of this paper is organized as follows.
Section~\ref{sec:assumption} describes the assumption and terminology
used this paper.
Section~\ref{sec:traffic_simulation} briefly explains our traffic
simulation program.
Section~\ref{sec:gpu_implementations} presents our schemes of GPU
implementations of traffic simulation.
Section~\ref{sec:evaluation} evaluates the benefit of our GPU
implementations, and provides an insight into future work.
Section~\ref{sec:conclusion} concludes this paper.

\section{Assumption and Terminology}
\label{sec:assumption}

We assume the Compute Unified Device Architecture (CUDA) for GPU
programming.
In CUDA, a unit of pieces of code that is launched on the GPU is called
a \textit{kernel}.
The kernel is typically composed of multiple parallelized
\textit{threads} that accelerate computations.
A unit of threads is called a \textit{block} where most computing
resources are shared.
A collection of blocks for the corresponding kernel is called a
\textit{grid}. 
The maximum number of threads that an individual block can contain is
defined by the GPU architecture.

GPU applications use a set of the API supported by CUDA.
We typically take the following steps to use the GPU: (i) allocate space
to device memory, (ii) copy data to the allocated device memory space,
(iii) launch the program on the GPU, (iv) copy resultant data back to
host memory, and (v) free the allocated device memory space.

@@

PLACEHOLDER.

@@

Figure~\ref{fig:sample-in-cpu} shows a brief overview of the
Sample-in-CPU scheme.
This scheme also has several alternatives depending on how many kernels
need to be launched on the GPU in each period. 
Suppose that we want to offload $n$ pieces of \textit{for} loops on to
the GPU.
We may return to the CPU $n$ times in totall, that is, return every time
one \textit{for} loop breaks, or otherwise we may just return once when
all the $n$ \textit{for} loops end.
This is a design decision, and is also dependent on the GPU architecture
and the parallelization structure.
In general, parallelized threads need to be synchronized when moving
across basic blocks.
Specifically, when moving to the next \textit{for} loop executed in
parallel, all the threads relevant to this parallel computing procedure
must synchronize with each other.
However, the maximum number of threads that can be synchronized on the
GPU is often limited.
As of 2012, for example, NVIDIA's GPU architectures limit the number of
such threads to 1024 or less~\cite{NVIDIA_Kepler}.
Our implementation therefore forces the program to return to the CPU
every time one \textit{for} loop breaks.
Note that this is not a conceptual limitation of GPU computing, but is a
current limitation of hardware.
We believe that this limitation would be removed or mitigated in
next-generation GPU architectures.

@@

We evaluate the performance of our GPU-accelerated traffic simulation
programs, using an NVIDIA GeForce GTX 560 Ti graphics card.
This is a middle-end graphics card based on the Fermi
architecture~\cite{NVIDIA_Fermi}, integrating 394 compute cores on a chip.
The programs used in this evaluation are written in CUDA, and are
compiled by the NVIDIA CUDA Compiler (NVCC) v4.2 suite~\cite{CUDA42}.

We focus on the one-dimensional OV model~\cite{Bando1995}, which can be
applied to freeways simulation.
The original program~\cite{OV_Program} of this simulation is written in
C, without any parallelization.
We use the same simulation parameters as the original default setup,
while increasing the number of agents (cars) to see how simulation
performance varies. 
The evaluation is conducted by comparing our GPU implementations with
the original CPU implementation.

\subsection{Performance Benefit}

\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{eps/eval_accel.eps}
\caption{Performance benefit of the GPU.}
\label{fig:eval_benefit}
\end{figure}

Figure~\ref{fig:eval_benefit} demonstrates the performance benefit of
the GPU under the current hardware limitation.
The Sample-in-CPU scheme herein exploits synchronization between the CPU
and the GPU after every \textit{for} loop to overcome the limitation that
CUDA threads cannot synchronize beyond blocks.
Despite of this restricted implementation, our GPU implementation
outperforms the original CPU implementation by about five times in
simulation time, when the number of agents exceeds a scale of 1000.
This evinces an affinity of the GPU and parallel computing for traffic
simulation.
It is also interesting to see that the CPU implementation has a better
performance for a small number of agents.
Thus, the overhead of communication between the CPU and the GPU has
non-trivial impact if the achievable parallelism is not sufficient.
Another notable observation is that the diffence in performance of the
CPU and the GPU implementations is saturated when the number of agents
reaches 1000.
This saturation implies the hardware potential of the NVIDIA GeForce GTX
560 Ti graphics card used in this evaluation.
Using high-end graphics cards, the difference in performance may not be
saturated at the same scale.

\subsection{Impact of Hardware Limitation}

\begin{figure}[t]
\centering
\includegraphics[width=0.5\textwidth]{eps/eval_nosync.eps}
\caption{Impact of the hardware limitation.}
\label{fig:eval_nosync}
\end{figure}

Figure~\ref{fig:eval_nosync} demonstrates the impact of the current
hardware limitation that CUDA threads cannot synchronize beyond blocks.
In this experiment, we observe ``what would happen if this hardware
limitation is removed?'', compromising the preciseness of simulation
results.
Namely, we use the current CUDA synchronization function to synchronize
threads running in different blocks, even though it does not work as
expected.
All the GPU implementations of the plotted schemes use this assumption.
Therefore, the Sample-in-CPU scheme is now different from what we have
evaluated in Figure~\ref{fig:eval_benefit}, since there is no need
anymore to return to the CPU every time one \textit{for} loop breaks.
We instead put a synchronization function between \textit{for} loops on
the GPU.
The Sample-in-CPU-Sync scheme is an alternative version of the
Sample-in-CPU scheme in that we download intermediate results from the
GPU to the CPU at the end of each sampling period.
The Sample-in-GPU scheme is what is presented in
Section~\ref{sec:sample-in-gpu}.

It is very notable that the Sample-in-GPU scheme improves the simulation
time by a factor of 1000 as compared to the original CPU implementation.
This result encourage suport for thread synchronization among blocks in
CUDA programming.
There are also several interesting observations obtained in this
experiment.
Comparing the Sample-in-CPU and the Sample-in-CPU-Sync schemes, it turns
out that the cost of downloading intermediate results from the GPU to
the CPU is becoming trivial as the scale of simulation increases.
Hence, a common argument of ``host-device data communication in GPU
programming is expensive'' does not apply to large-scale traffic
simulation.
Another issue of concern often discussed in GPU programming is that
a single-thread performance of GPUs is weak as compared to that of CPUs.
This is true in general, but our experimental result shows that the
performance loss caused by pushing most pieces of code into the GPU in
the Sample-in-GPU scheme is not signficant even for a small number of
agents.
This observation leads to some conclusion that the Sample-in-GPU scheme
would be the best choice for any scale of traffic simulation.

\subsection{Practical Concern}

@@

We finally examine the impact of misbehavior of thread synchronization
in our GPU implementations.
The simulation results produced by our GPU implementations are not
precise except for those of the Sample-in-CPU scheme presented in
Figure~\ref{fig:eval_benefit}, due to our optimistic use of thread
synchronization on the GPU.
We need to wait for new GPU architectures removing the limitation of
thread synchronization in order to make them precise.
Practically speaking, however, imprecise results of simulation are still
valuable and meaningful as far as they are limited to an acceptable
range of errors for traffic flow.

Figure~\ref{fig:eval_error} demonstrates the error percentage of
location and velocity between the simulation results produced by the
original CPU and the imprecise Sample-in-GPU implementations.
Since this is one-dimensional OV simulation, there are only forward or
backward in location, and high or low in velocity.
Under this constraint, the error in location is very trivial.
This is because the characteristic of traffic flow is well captured,
even though the simulated locations and velocities have some error in
their individual values.
A degree of the error in velocity is slightly high as compared to that
in location.
This means that velocity is more sensitive to imprecise results than
location.
However, the maximum error in velocity observed in our experiment is at
most 5\%.
Suppose that the average driving speed is 50mph.
The simulated speed in our imprecise implementation would still be
within a range of 47.5 to 52.5mph.
Intuitively, this scale of error is acceptable in practice.
We plan to further investigate if the same argument can be applied to a
more variety of parameter setups and multidimensional OV simulation
programs.

@@

PLACEHOLDER.

Imprecise Computation~\cite{Lin1987}.
Gdev~\cite{Kato2012}.
